The Second Data Release of the INT Photometric H$\alpha$ Survey of the Northern Galactic Plane (IPHAS DR2)

### 2014-03-17 paper/iphas-dr2.tex
Various minor improvements

\author[G. Barentsen
et. al]{G. Barentsen$^{1}$\thanks{E-mail:geert@barentsen.be},
H. J. Farnhill$^1$,
J. E. Drew$^1$,
E. A. Gonz$\acute{\rm{a}}$lez-Solares$^2$, 
R. Greimel$^3$, \newauthor
M. J. Irwin$^2$,
B. Mizalski$^{4}$,
C. Ruhland$^1$,
P. Groot$^5$,
A. Mampaso$^6$, 
S. E. Sale$^7$, \newauthor
M. J. Barlow$^{8}$,
R. L. M Corradi$^{6}$,
J. J. Drake$^{9}$,
J. Fabregat$^{10}$, 
B. T. Gaensicke$^{11}$, \newauthor
A. S. Hales$^{12}$,
J. Irwin$^{9}$,
C. Knigge$^{13}$,
D. J. Lennon$^{14}$, 
J. R. Lewis$^{2}$,
M. Mohr-Smith$^{1}$, \newauthor
R. A. H. Morris$^{15}$,
T. Naylor$^{16}$,
Q. A. Parker$^{17}$, 
S. Phillipps$^{13}$,
R. Raddi$^{11}$, \newauthor
P. Rodriguez-Gil$^{6}$,
L. Sabin$^{18}$,
D. Steeghs$^{11}$, 
Y. C. Unruh$^{19}$,
K. Viironen$^{20}$, \newauthor
J. S. Vink$^{21}$, 
N. A. Walton$^{2}$,
N. J. Wright$^{1}$,
A. A. Zijlstra$^{22}$.
\newauthor\\
$^{1}$School of Physics, Astronomy \& Mathematics, University of Hertfordshire, College Lane, Hatfield, Hertfordshire, AL10 9AB, U.K.\\
$^{2}$Institute of Astronomy, Cambridge University, Madingley Road, Cambridge, CB3 OHA, U.K.\\
$^{3}$IGAM, Institute of Physics, University of Graz, Universit\"atsplatz 5, Graz, Austria.\\
$^{4}$South African Astronomical Observatory, P.O. Box 9, Observatory, 7935 Cape Town, South Africa.\\
$^{5}$Afdeling Sterrenkunde, Radboud Universiteit Nijmegen, Faculteit NWI, Postbus 9010, 6500 GL Nijmegen, The Netherlands.\\
$^{6}$Instituto de Astrof\'isica de Canarias, 38200 La Laguna, Tenerife, Spain.\\
$^{7}$Rudolf Peierls Centre for Theoretical Physics, Keble Road, Oxford, OX1 3NP, U.K.\\
$^{8}$University College London, Department of Physics \& Astronomy, 
Gower Street, London WC1E 6BT, U.K.\\
$^{9}$Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, 
Cambridge, MA 02138, U.S.A. \\
$^{10}$Observatorio Astr\'onomico, Universidad de Valencia,
Catedr\'atico Jos\'e Beltr\'an 2, 46980 Paterna, Spain\\
$^{11}$Department of Physics, University of Warwick, Gibbet Hill Road, Coventry, CV4 7AL, U.K.\\
$^{12}$ Joint ALMA Observatory, Alonso de Córdova 3107, Vitacura 763-0355, Santiago, Chile. \\
$^{13}$School of Physics \& Astronomy, University of Southampton, Southampton, SO17 1BJ, U.K.\\
$^{14}$European Space Astronomy Centre (ESAC), Villafranca del Castillo, Villanueva de la Canada, E-28692 Madrid, Spain.\\
$^{15}$School of Physics, Bristol University, Tyndall Avenue, Bristol, BS8 1TL, U.K.\\
$^{16}$School of Physics, University of Exeter, Stocker Road, Exeter, EX4 4QL, U.K.\\
$^{17}$Department of Physics \& Astronomy, Macquarie University, NSW 2109, Australia\\
$^{18}$Instituto de Astonom\'ia y Meteorolog\'ia, Departamento de F\'isica, CUCEI, Universidad de Guadalajara, Mexico.\\
$^{19}$Department of Physics, Blackett Laboratory, Imperial College
London, Prince Consort Road, London, SW7 2AZ, U.K.\\
$^{20}$Centro de Estudios de F\'isica del Cosmos de Arag\'on, Plaza San Juan 1, Planta 2, Teruel, 44001, Spain. \\
$^{21}$Armagh Observatory, College Hill, Armagh, Northern Ireland,
BT61 9DG, U.K.\\
$^{22}$Jodrell Bank Centre for Astrophysics, School of Physics \&
Astronomy, University of Manchester, Oxford Road, Manchester M13 9PL, U.K.}

@@

\begin{abstract} % BACKGROUND, OBJECTIVE, METHODS, RESULTS, CONCLUSIONS
The INT/WFC Photometric H$\alpha$ Survey 
of the Northern Galactic Plane (IPHAS)
is a 1800 deg$^2$ imaging survey
covering the entire northern Milky Way at $|b| < 5^\circ$
in the $r$, $i$ and \ha\ filters 
using the Wide Field Camera (WFC) 
on the 2.5-metre Isaac Newton Telescope (INT) in La Palma.
We present the first quality-controlled and
uniformly-calibrated source catalogue
to have been extracted from the images,
providing single-epoch photometry
for 219 million unique sources
across 92\% of the survey area.
The observations were carried out between 2003 and 2013
at a median seeing of 1.1 arcsec
and to a mean depth of $r=21.2$ / $i=20.0$ / $\ha=20.3$
($5\sigma$ limits, Vega system).
We explain the data reduction 
and quality control procedures,
describe and test the new uniform calibration,
and detail the construction of the 97-column source catalogue.
We show that the global calibration is accurate to
0.03~mag (rms) by validating against SDSS,
and provide a series of recommended quality criteria
which can be used to select the most reliable data
from the catalogue.
Finally, we demonstrate the ability of the 
catalogue's unique
$(r-i)$/($r-\ha)$ colour-colour diagram to
(i) characterise stellar populations and extinction regimes
towards different Galactic sight-lines
and (ii) select candidate \ha\ emission-line objects.
IPHAS is the first survey to offer comprehensive CCD photometry
of point sources across the Galactic Plane at visible wavelengths,
providing the much-needed counterpart
to infrared surveys.
We also provide images for some of the most
crowded regions to be faced by Gaia.
\end{abstract}

@@

An initial list of candidate \ha-emitters
based on the first IPHAS data release was previously
presented by \cite{Witham2008}. 
Because no uniform calibration was available
at the time, \citeauthor{Witham2008} employed 
a sigma-clipping technique to select objects with
large, outlying $r$-\ha\ colours.
In contrast, the new catalogue
allows objects to be picked out
from the $(r-i)$/$(r-\ha)$ colour-colour diagram
using model-based colour criteria
rather than a statistical procedure.
In what follows we demonstrate this ability 
by selecting candidate emission-line objects
towards a small region in the sky.

@@

Figure~\ref{fig:mosaic_iphas} shows a 20-by-15 arcmin
colour mosaic centred on Sh 2-82,
composed of our \ha\ (red channel),
$r$ (green channel),
and $i$ (blue channel) images.
The ionising star can be seen as the bright object
in the centre of the H{\sc ii} region,
which is surrounded by a faint reflection nebula
and several dark cloud filaments.
For comparison, Figure~\ref{fig:mosaic_spitzer} shows
a mosaic of the same region 
as seen in the mid-infrared by the Spitzer Space Telescope
The infrared image reveals a bubble-shaped structure
of warm dust which surrounds the entire H{\sc ii} region,
and is thought to contain a wealth of infrared-bright
young stellar objects \citep{Yu2012}.
Many of these young objects 
appear as red- and pink-coloured stars in the Spitzer mosaic --
located predominantly in the north-western part
of the bubble-shaped structure.

\begin{figure*}
    \begin{minipage}[b]{0.8\linewidth}
        \includegraphics[width=\textwidth]{./plots/mosaic/sh2-82-iphas.pdf} 
    \end{minipage}
\caption{IPHAS-based mosaic of H{\sc ii} region Sh 2-82,
composed of \ha\ (red channel), $r$ (green channel) and $i$ (blue channel). Yellow triangles show the position of candidate \ha-emitters
which have been selected from the colour-colour diagram
in Figure~\ref{fig:emitters}. Note that the H{\sc ii} region is surrounded by a faint blue/green reflection nebula
and dark cloud filaments.}
\label{fig:mosaic_iphas}
    \begin{minipage}[b]{0.8\linewidth}
        \includegraphics[width=\textwidth]{./plots/mosaic/sh2-82-spitzer.pdf} 
    \end{minipage}
    \caption{Star-forming region Sh 2-82 as seen in the mid-infrared
    by the Spitzer Space Telescope. The mosaic is composed of the 24\,\micron\ (red), 8.0\,\micron\ (green) and 4.5\,\micron\ (blue) bands.
    The image reveals a bubble-shaped structure which surrounds the {\sc Hii} region that is seen in the IPHAS mosaic of the same region (Figure~\ref{fig:mosaic_iphas}).}
    \label{fig:mosaic_spitzer}
\end{figure*}

Figure~\ref{fig:emitters} presents
the IPHAS colour-colour diagram for 
the 20-by-15 arcmin region shown in the mosaics.
Gray circles show all objects
which are brighter than $r<20$
and have been flagged as \emph{reliable}
in our catalogue.
The diagram also shows the unreddened main sequence (solid line)
and the expected position of unreddened main-sequence stars
with \ha\ in emission
at a strength of EW=$-10{\rm \AA}$ (dashed line),
taken from the colour simulations due to \citet{Barentsen2011a}.
Six stars are found to lie confidently above the 
dashed line at the level of $3\sigma$,
that is, the distance between the objects and the dashed line
is larger than three times the uncertainty on the $r$-\ha\ colour.
These reliable candidate \ha-emitters
are marked by red triangles in the colour-colour diagram
and their photometry is detailed in Table~\ref{tbl:emitters}.
The spatial distribution of the six candidate emission-line objects
is marked by yellow triangles in the colour mosaic (Fig.~\ref{fig:mosaic_iphas}).

@@

In previous work, we have shown that the majority of
\ha-emitters seen towards an H{\sc ii} region
are likely to be Classical T Tauri Stars \citep{Barentsen2011a}.
These are young objects which show \ha\ in emission
due to the presence of hot, infalling gas
which is accreting onto the star
from a circumstellar disk.
This is likely to be the case for the candidate \ha-emitters
we discovered towards Sh 2-82 as well.
Two of our candidates have recently been identified
as Young Stellar Objects (YSO)
in an initial investigation of the region by \cite{Yu2012}.
In their study, the authors used 2MASS and Spitzer data
to select likely young stars by selecting objects with
an infrared colour excess, consistent with 
the presence of a heated circumstellar disk.
Although \citeauthor{Yu2012} did not identify
four of our candidates as having
an infrared excess, we note that each of these four objects
is detected at S/N$>$? in the Spitzer 8.0\,\micron\ image,
which is consistent with an optically-unveiled Class II YSO.

%The IPHAS catalogue offers photometry
%for 2197 objects in the region down to $r<20$.

Sh 2-82 is one of a large population of poorly-studied star-forming regions
located at low Galactic latitudes,
which have only recently started to become revealed
by efforts to catalogue the wealth of ``bubbles'' detected
by mid-infrared surveys \cite[e.g.][]{Churchwell2006,Simpson2012},
and by efforts to catalogue previously unknown clusters seen 
at near-infrared wavelengths \cite[e.g.][]{Bica2003}.
The optical data provided by IPHAS offers a handle -- or lower limit --
on the extinction and distance towards many of these regions.
When the extinction is sufficiently low,
we have shown here that IPHAS can discover optically-unveiled 
members of the region.

%TODO: annotate -10 EW text on plot?
%TODO: mention Gaia, e.g. IPHAS for transient identification?
%\subsection{Caveats and lessons learnt}
%\subsection{Comparison with Gaia}?

@@

The catalogue is made available through the Vizier
catalogue tool (http://vizier.u-strasbg.fr),
where it is known as the ``IPHAS DR2 Source Catalogue''
and where it can be queried
(catalogue ID ??? Conesearch/TAP URL??).
The catalogue can also be downloaded in its entirety
from the IPHAS website (www.iphas.org) as a collection 
of binary FITS tables which comprise 50\,GB,
together with a script
to ingest these tables into a PostgreSQL database.
Finally, the website also provides access to the pipeline-processed
imaging data, which we have updated to include
the re-calibrated zeropoints and the best-available
astrometric solution in the image headers.

@@

In future work, we intend to draw upon the PanSTARRS photometric
survey to further improve the accuracy of our global calibration.
We will also aim to correct the IPHAS photometry for the
radial field distortions,
and plan to produce a catalogue 
which details all the secondary detections
to aid time-domain studies.

@@

The IPHAS survey was carried out 
at the Isaac Newton Telescope (INT).
The INT is operated on the island of La Palma
by the Isaac Newton Group
in the Spanish Observatorio del Roque de los Muchachos
of the Instituto de Astrofisica de Canarias.
All data were processed 
by the Cambridge Astronomical Survey Unit
at the Institute of Astronomy in Cambridge.

@@

This research made extensive use of
several complementary photometric surveys.
Our global calibration was aided
by the AAVSO Photometric All-Sky Survey (APASS),
funded by the Robert Martin Ayers Sciences Fund.
The calibration was tested against the
Sloan Digitized Sky Survey (SDSS),
funded by the Alfred P. Sloan Foundation,
the Participating Institutions,
the National Science Foundation,
the U.S. Department of Energy,
the National Aeronautics and Space Administration,
the Japanese Monbukagakusho, the Max Planck Society,
and the Higher Education Funding Council for England.
The astrometric pipeline reduction made
significant use of the Two Micron All Sky Survey (2MASS),
which is a joint project 
of the University of Massachusetts
and the Infrared Processing and Analysis Center/
California Institute of Technology,
funded by NASA and the NSF.

>>>>

\author[G. Barentsen
et. al]{G. Barentsen$^{1}$\thanks{E-mail:geert@barentsen.be},
H. J. Farnhill$^1$,
J. E. Drew$^1$,
E. A. Gonz$\acute{\rm{a}}$lez-Solares$^2$, 
R. Greimel$^3$, \newauthor
M. J. Irwin$^2$,
B. Mizalski$^{4}$,
C. Ruhland$^1$,
P. Groot$^5$,
A. Mampaso$^6$, 
S. E. Sale$^7$, \newauthor
M. J. Barlow$^{8}$,
R. L. M Corradi$^{6}$,
J. J. Drake$^{9}$,
J. Fabregat$^{10}$, 
B. T. Gaensicke$^{11}$, \newauthor
A. S. Hales$^{12}$,
J. Irwin$^{9}$,
C. Knigge$^{13}$,
T. Kupfer$^{5}$,
D. J. Lennon$^{14}$, 
J. R. Lewis$^{2}$, \newauthor
M. Mohr-Smith$^{1}$, 
R. A. H. Morris$^{15}$,
T. Naylor$^{16}$,
Q. A. Parker$^{17}$, 
S. Phillipps$^{13}$, \newauthor
R. Raddi$^{11}$, 
P. Rodriguez-Gil$^{6}$,
L. Sabin$^{18}$,
D. Steeghs$^{11}$, 
Y. C. Unruh$^{19}$, \newauthor
K. Viironen$^{20}$, 
J. S. Vink$^{21}$, 
N. A. Walton$^{2}$,
N. J. Wright$^{1}$,
A. A. Zijlstra$^{22}$.
\newauthor\\
$^{1}$School of Physics, Astronomy \& Mathematics, University of Hertfordshire, College Lane, Hatfield, Hertfordshire, AL10 9AB, U.K.\\
$^{2}$Institute of Astronomy, Cambridge University, Madingley Road, Cambridge, CB3 OHA, U.K.\\
$^{3}$IGAM, Institute of Physics, University of Graz, Universit\"atsplatz 5, Graz, Austria.\\
$^{4}$South African Astronomical Observatory, P.O. Box 9, Observatory, 7935 Cape Town, South Africa.\\
$^{5}$Afdeling Sterrenkunde, Radboud Universiteit Nijmegen, Faculteit NWI, Postbus 9010, 6500 GL Nijmegen, The Netherlands.\\
$^{6}$Instituto de Astrof\'isica de Canarias, 38200 La Laguna, Tenerife, Spain.\\
$^{7}$Rudolf Peierls Centre for Theoretical Physics, Keble Road, Oxford, OX1 3NP, U.K.\\
$^{8}$University College London, Department of Physics \& Astronomy, 
Gower Street, London WC1E 6BT, U.K.\\
$^{9}$Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, 
Cambridge, MA 02138, U.S.A. \\
$^{10}$Observatorio Astr\'onomico, Universidad de Valencia,
Catedr\'atico Jos\'e Beltr\'an 2, 46980 Paterna, Spain\\
$^{11}$Department of Physics, University of Warwick, Gibbet Hill Road, Coventry, CV4 7AL, U.K.\\
$^{12}$ Joint ALMA Observatory, Alonso de Córdova 3107, Vitacura 763-0355, Santiago, Chile. \\
$^{13}$School of Physics \& Astronomy, University of Southampton, Southampton, SO17 1BJ, U.K.\\
$^{14}$European Space Astronomy Centre (ESAC), Villafranca del Castillo, Villanueva de la Canada, E-28692 Madrid, Spain.\\
$^{15}$School of Physics, Bristol University, Tyndall Avenue, Bristol, BS8 1TL, U.K.\\
$^{16}$School of Physics, University of Exeter, Stocker Road, Exeter, EX4 4QL, U.K.\\
$^{17}$Department of Physics \& Astronomy, Macquarie University, NSW 2109, Australia\\
$^{18}$Instituto de Astonom\'ia y Meteorolog\'ia, Departamento de F\'isica, CUCEI, Universidad de Guadalajara, Mexico.\\
$^{19}$Department of Physics, Blackett Laboratory, Imperial College
London, Prince Consort Road, London, SW7 2AZ, U.K.\\
$^{20}$Centro de Estudios de F\'isica del Cosmos de Arag\'on, Plaza San Juan 1, Planta 2, Teruel, 44001, Spain. \\
$^{21}$Armagh Observatory, College Hill, Armagh, Northern Ireland,
BT61 9DG, U.K.\\
$^{22}$Jodrell Bank Centre for Astrophysics, School of Physics \&
Astronomy, University of Manchester, Manchester M13 9PL, U.K.}

@@

\begin{abstract} % BACKGROUND, OBJECTIVE, METHODS, RESULTS, CONCLUSIONS
The INT/WFC Photometric H$\alpha$ Survey 
of the Northern Galactic Plane (IPHAS)
is a 1800 deg$^2$ imaging survey
covering the entire northern Milky Way at $|b| < 5^\circ$
in the $r$, $i$ and \ha\ filters 
using the Wide Field Camera (WFC) 
on the 2.5-metre Isaac Newton Telescope (INT) in La Palma.
We present the first quality-controlled and
uniformly-calibrated source catalogue
to have been extracted from the images,
providing single-epoch photometry
for 219 million unique sources
across 92\% of the survey area.
The observations were carried out between 2003 and 2013
at a median seeing of 1.1 arcsec
and to a mean depth of $r=21.2$ / $i=20.0$ / $\ha=20.3$
($5\sigma$ limits, Vega system).
We explain the data reduction 
and quality control procedures,
describe and test the new uniform calibration,
and detail the construction of the 97-column source catalogue.
We show that the global calibration is accurate to
0.03~mag (rms) by validating against SDSS,
and provide a series of recommended quality criteria
which can be used to select the most reliable data
from the catalogue.
Finally, we demonstrate the ability of the 
catalogue's unique
$(r-i)$/($r-\ha)$ colour-colour diagram to
(i) characterise stellar populations and extinction regimes
towards different Galactic sight-lines
and (ii) select candidate \ha\ emission-line objects.
IPHAS is the first survey to offer comprehensive CCD photometry
of point sources across the Galactic Plane at visible wavelengths,
providing the much-needed counterpart
to infrared surveys.
IPHAS also provide images for some of the most
crowded regions to be faced by Gaia.
\end{abstract}

@@

An initial list of candidate \ha-emitters
based on the first IPHAS data release was previously
presented by \cite{Witham2008}. 
Because no uniform calibration was available
at the time, \citeauthor{Witham2008} employed 
a sigma-clipping technique to select objects with
large, outlying $(r-\ha)$ colours.
In contrast, the new catalogue
allows objects to be picked out
from the $(r-i)$/$(r-\ha)$ colour-colour diagram
using model-based colour criteria
rather than a statistical procedure.
In what follows we demonstrate this ability 
by selecting candidate emission-line objects
towards a small region in the sky.

@@

Figure~\ref{fig:mosaic_iphas} shows a 20-by-15 arcmin
colour mosaic centred on Sh 2-82,
composed of our \ha\ (red channel),
$r$ (green channel),
and $i$ (blue channel) images.
The ionising star can be seen as the bright object
in the centre of the H{\sc ii} region,
which is surrounded by a faint reflection nebula
and several dark cloud filaments.
For comparison, Figure~\ref{fig:mosaic_spitzer} shows
a mosaic of the same region 
as seen in the mid-infrared by the Spitzer Space Telescope's
Galactic Plane survey \citep[GLIMPSE;][]{Benjamin2003,Churchwell2009}.
The infrared image reveals a bubble-shaped structure
which is thought to originate from the
mid-infrared emission of Poly-Aromatic Hydrocarbons (PAHs)
-- i.e. warm dust --
which is frequently observed
at the interface between neutral regions of interstellar material
and the ionising radiation from early-type stars \citep{Churchwell2006}.
In this example, \cite{Yu2012}
recently noted that the warm dust
that surrounds Sh~2-82 
appears to contain infrared-bright
Young Stellar Objects (YSOs).
Many of these young objects 
appear as red- and pink-coloured stars
in Fig.~\ref{fig:mosaic_spitzer},
located predominantly in the top-left part
of the bubble.

\begin{figure*}
    \begin{minipage}[b]{0.8\linewidth}
        \includegraphics[width=\textwidth]{./plots/mosaic/sh2-82-iphas.pdf} 
    \end{minipage}
\caption{IPHAS-based mosaic of H{\sc ii} region Sh 2-82,
composed of \ha\ (red channel), $r$ (green channel) and $i$ (blue channel). Yellow triangles show the position of candidate \ha-emitters
which have been selected from the colour-colour diagram
in Figure~\ref{fig:emitters}. Note that the H{\sc ii} region is surrounded by a faint blue/green reflection nebula
and dark cloud filaments.}
\label{fig:mosaic_iphas}
    \begin{minipage}[b]{0.8\linewidth}
        \includegraphics[width=\textwidth]{./plots/mosaic/sh2-82-spitzer.pdf} 
    \end{minipage}
    \caption{Star-forming region Sh 2-82 as seen in the mid-infrared
    by the Spitzer Space Telescope's GLIMPSE survey. The mosaic is composed of the 24\,\micron\ (red), 8.0\,\micron\ (green) and 4.5\,\micron\ (blue) bands.
    The image reveals a bubble-shaped structure which surrounds the {\sc Hii} region that is seen in the IPHAS mosaic of the same region (Figure~\ref{fig:mosaic_iphas}). 
    This structure has previously been labelled as N115 in the 
catalogue due to \citet{Churchwell2006}.}
    \label{fig:mosaic_spitzer}
\end{figure*}

Figure~\ref{fig:emitters} presents
the IPHAS colour-colour diagram for 
the 20-by-15 arcmin region shown in the mosaics.
Gray circles show all objects
which are brighter than $r<20$
and have been flagged as \emph{reliable}
in our catalogue.
The diagram also shows the unreddened main sequence (solid line)
and the expected position of unreddened main-sequence stars
with \ha\ in emission
at a strength of EW=$-10{\rm \AA}$ (dashed line),
taken from the colour simulations due to \citet{Barentsen2011a}.
Six stars are found to lie confidently above the 
dashed line at the level of $3\sigma$,
that is, the distance between the objects and the dashed line
is larger than three times the uncertainty
in their $(r-\ha)$ colour.
These reliable candidate \ha-emitters
are marked by red triangles in the colour-colour diagram
and their photometry is detailed in Table~\ref{tbl:emitters}.
The spatial distribution of the six candidate emission-line objects
is marked by yellow triangles in the colour mosaic (Fig.~\ref{fig:mosaic_iphas}).

@@

In previous work, we have shown that the majority of
\ha-emitters seen towards an H{\sc ii} region
are likely to be Classical T Tauri Stars \citep{Barentsen2011a}.
These are young objects
which are thought to show \ha\ in emission
due to the presence of hot, infalling gas
which is accreting onto the star
from a circumstellar disk.
This is likely to be the case for the candidate \ha-emitters
we discovered towards Sh 2-82 as well.
Two of our candidates have recently been identified
as YSOs in the initial investigation of the region by \cite{Yu2012}.
In their study, the authors used 2MASS and Spitzer data
to select likely young stars by selecting objects with
an infrared colour excess, consistent with 
the presence of a heated circumstellar disk.
Although \citeauthor{Yu2012} did not identify
the other four candidate emitters in our sample
as having an infrared excess,
we note that each of these four objects
are detected in the Spitzer 8.0\,\micron\ image
at S/N$>$5,
which is consistent with the objects
being optically-unveiled Class II YSOs
with a mild infrared excess.

Sh 2-82 is one of a large population of poorly-studied star-forming regions
located at low Galactic latitudes,
which have only recently started to become revealed
by efforts to catalogue the wealth of ``bubbles'' detected
at mid-infrared wavelengths \citep{Churchwell2006,Simpson2012},
and by efforts to catalogue previously unknown clusters seen 
in the near-infrared \cite[e.g.][]{Bica2003}.
The data provided by IPHAS offers a handle
or lower limit on the extinction and distance
towards these regions,
and can be used to discover optically-unveiled
T~Tauri stars when the extinction is sufficiently low.

@@

The catalogue is made available through the Vizier
catalogue tool (http://vizier.u-strasbg.fr),
where it is known as the ``IPHAS DR2 Source Catalogue''
and where it can be queried
(catalogue ID TBD; Conesearch/TAP URL TBD?).
The catalogue can also be downloaded in its entirety
from the IPHAS website (www.iphas.org) as a collection 
of binary FITS tables which comprise 50\,GB,
together with a script
to ingest these tables into a PostgreSQL database.
Finally, the website also provides access to the pipeline-processed
imaging data, which we have updated to include
the re-calibrated zeropoints and the best-available
astrometric solution in the image headers.

@@

In future work, we intend to draw upon the PanSTARRS photometric
survey to further improve the accuracy of our global calibration.
We also aim to correct the IPHAS photometry for the
radial field distortions,
plan to produce a catalogue 
which details all the secondary detections
to aid time-domain studies,
and are investigating the use of photometry
derived from PSF-fitting across the survey.
%TODO: mention Gaia, e.g. IPHAS for transient identification?

@@

The IPHAS survey was carried out 
at the Isaac Newton Telescope (INT).
The INT is operated on the island of La Palma
by the Isaac Newton Group
in the Spanish Observatorio del Roque de los Muchachos
of the Instituto de Astrofisica de Canarias.
We are deeply indebted to the ING staff and students
for their ongoing support of the INT.
All data were processed 
by the Cambridge Astronomical Survey Unit
at the Institute of Astronomy in Cambridge.

@@

This research made extensive use of
several complementary photometric surveys.
Our global calibration was aided
by the AAVSO Photometric All-Sky Survey (APASS),
funded by the Robert Martin Ayers Sciences Fund.
The calibration was tested against the
Sloan Digitized Sky Survey (SDSS),
funded by the Alfred P. Sloan Foundation,
the Participating Institutions,
the National Science Foundation,
the U.S. Department of Energy,
the National Aeronautics and Space Administration,
the Japanese Monbukagakusho, the Max Planck Society,
and the Higher Education Funding Council for England.
The astrometric pipeline reduction made
significant use of the Two Micron All Sky Survey (2MASS),
which is a joint project 
of the University of Massachusetts
and the Infrared Processing and Analysis Center/
California Institute of Technology,
funded by NASA and the NSF.
This work includes observations made
with the Spitzer Space Telescope,
which is operated by the Jet Propulsion Laboratory,
California Institute of Technology under a contract with NASA.
